---
title: "Air Pollution in Berlin during Corona"
author: "Maximilian Nölscher"
date: "`r format(Sys.Date())`"
output: github_document
---
```{r, include=FALSE}
purrr::map(
  list.files(
    path = "./functions",
    pattern = "\\.R$",
    full.names = TRUE,
    recursive = TRUE
  ),
  source
)
```

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.align = "center")
knitr::opts_chunk$set(fig.width = 10)
```

```{r}
library(sf)
library(tidyverse)
library(lubridate)
```


# Data Import

Path to file
```{r}
path_to_file <- "raw_data/verkehr/"
```

List all files in path
```{r}
files <- path_to_file %>%
  list.files() %>%
  paste0(path_to_file, .)
```

```{r}
if (str_detect(path_to_file, "verkehr")) {
  data_background <- files %>%
    map_dfr(~ read_csv2(., skip = 3))
} else if (str_detect(path_to_file, "hintergrund")) {
  data_background <- files %>%
    map(~ read_csv2(., skip = 3)) %>%
    map(~ select(., 1:5)) %>%
    bind_rows()
}
```

Column names
```{r}
if (str_detect(path_to_file, "verkehr")) {
  col_names <- files[1] %>%
    read_csv2() %>%
    slice(1) %>%
    pivot_longer(cols = everything()) %>%
    pull(value)
} else if (str_detect(path_to_file, "hintergrund")) {
  col_names <- files[1] %>%
    read_csv2() %>%
    slice(1) %>%
    pivot_longer(cols = everything()) %>%
    slice(1:5) %>%
    pull(value)
}
```

Rename columns of `data_background`
```{r}
data_background <- data_background %>%
  set_names(col_names) %>%
  janitor::clean_names() %>%
  rename(date = messkomponente)
```

Fix date column
```{r}
data_background <- data_background %>%
  mutate(date = dmy_hm(date))
```

Remove duplicate observations
```{r}
data_background <- data_background %>%
  distinct(date, .keep_all = TRUE)
```


Add columns for year, month, ...
```{r}
data_background <- data_background %>%
  mutate(year = year(date)) %>%
  mutate(month = month(date)) %>%
  mutate(week = week(date)) %>%
  mutate(day = day(date)) %>%
  mutate(weekday = weekdays(date)) %>%
  mutate(daytime = hour(date)) %>%
  mutate(yearday = yday(date)) %>%
  select(date, year, month, week, day, weekday, daytime, yearday, everything())
```

Show the dataframe
```{r}
data_background
```

# First visualizations
```{r}
variables <- c("stickstoffmonoxid", "stickstoffdioxid", "stickoxide")
```

Clean outliers
```{r}
data_background <- data_background %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name) %>%
  mutate(value = ifelse(value >= 1e4, NA, value)) %>%
  mutate(value = zoo::na.approx(value)) %>%
  pivot_wider()
```


```{r}
data_background %>%
  pivot_longer(cols = one_of(variables)) %>%
  ggplot(aes(date, value, group = name)) +
  geom_line(
    colour = "red",
    alpha = .6
  ) +
  scale_x_datetime(
    date_breaks = "3 months",
    # date_minor_breaks = "1 month",
    date_labels = "%b\n'%y"
  ) +
  facet_wrap(~name, ncol = 1, scales = "free_y")
```

Weekly mean concentration
```{r}
data_background %>%
  select(-stickoxide) %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name, week) %>%
  summarise(mean = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(week, mean, group = name, fill = mean)) +
  geom_col(alpha = .6) +
  # scale_x_datetime(date_breaks = "month") +
  scale_fill_gradient(low = "darkgreen", high = "red") +
  scale_x_continuous(breaks = seq(0, 52, 5)) +
  facet_wrap(~name, ncol = 1, scales = "free_y")
```


Daily values compared to period mean
```{r}
data_background %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name) %>%
  mutate(mean = mean(value, na.rm = TRUE)) %>%
  group_by(name) %>%
  mutate(diff_to_mean = value - mean) %>%
  group_by(name, year, month, day) %>%
  summarise(daily_mean_diff_to_mean = mean(diff_to_mean, na.rm = TRUE)) %>%
  mutate(
    date = as_date(str_c(year, month, day, sep = "-")),
    colour_category = if_else(daily_mean_diff_to_mean <= 0, "negativ", "positiv")
  ) %>%
  ggplot(aes(date,
    daily_mean_diff_to_mean,
    group = name,
    fill = daily_mean_diff_to_mean
  )) +
  geom_col() +
  # scale_fill_manual() +
  scale_fill_gradient2(low = "darkgreen", mid = "snow3", high = "red") +
  scale_x_date(
    date_breaks = "3 month",
    date_minor_breaks = "1 month",
    date_labels = "%b\n'%y"
  ) +
  theme(legend.position = "none") +
  facet_wrap(~name, ncol = 1, scales = "free_y")
```

Mean daily values compared to period mean
```{r}
data_background %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name) %>%
  mutate(mean = mean(value, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(name) %>%
  mutate(diff_to_mean = value - mean) %>%
  ungroup() %>%
  group_by(name, year, month, day) %>%
  summarise(daily_mean_diff_to_mean = mean(diff_to_mean, na.rm = TRUE)) %>%
  mutate(
    date = as_date(str_c(year, month, day, sep = "-")),
    colour_category = if_else(daily_mean_diff_to_mean <= 0, "negative", "positive")
  ) %>%
  ggplot(aes(
    date,
    daily_mean_diff_to_mean
  )) +
  geom_area(
    fill = "red",
    alpha = .3
  ) +
  geom_ribbon(aes(ymin = 0, ymax = ifelse(daily_mean_diff_to_mean >= 0, 0, daily_mean_diff_to_mean)),
    fill = "green",
    alpha = .3
  ) +
  # scale_fill_manual() +
  # scale_fill_gradient2(low='darkgreen', mid='snow3', high='red') +
  scale_x_date(
    date_breaks = "3 month",
    date_minor_breaks = "1 month",
    date_labels = "%b\n'%y"
  ) +
  theme(legend.position = "none") +
  facet_wrap(~name, ncol = 1, scales = "free_y")
```

```{r}
days_in_week <- tibble(
  weekday = weekdays(x = as.Date(seq(7), origin = "1950-01-01")),
  weekday_id = 1:7
)
```

```{r}
days_in_week_vec <- days_in_week %>% pull(weekday_id)

names(days_in_week_vec) <- days_in_week %>% pull(weekday)
```

Mean concentrations per weekday
```{r}
data_background %>%
  select(-stickoxide) %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name, weekday) %>%
  summarise(mean_weekday = mean(value, na.rm = TRUE)) %>%
  group_by(name) %>%
  left_join(days_in_week) %>%
  arrange(weekday_id) %>%
  ggplot(aes(weekday_id, mean_weekday, fill = mean_weekday, group = name)) +
  geom_col(alpha = .5) +
  theme(legend.position = "none") +
  scale_x_continuous(
    breaks = 1:7,
    labels = pull(days_in_week, weekday)
  ) +
  scale_fill_gradient(low = "darkgreen", high = "red") +
  facet_wrap(~name, ncol = 1)
```

Mean concentrations per per hour per weekday
```{r}
data_background %>%
  select(-stickoxide) %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name, weekday, daytime) %>%
  summarise(mean_weekday_daytime = mean(value, na.rm = TRUE)) %>%
  group_by(name) %>%
  left_join(days_in_week, by = "weekday") %>%
  arrange(weekday_id, daytime) %>%
  ggplot(aes(daytime, mean_weekday_daytime, fill = mean_weekday_daytime, group = name)) +
  geom_col(alpha = .5) +
  theme(legend.position = "none") +
  scale_fill_gradient(low = "darkgreen", high = "red") +
  facet_grid(weekday_id ~ name, scales = "free_y")
```

Mean concentrations per per hour per weekday
```{r}
data_background %>%
  select(-stickoxide) %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name, weekday, daytime) %>%
  summarise(mean_weekday_daytime = mean(value, na.rm = TRUE)) %>%
  group_by(name) %>%
  left_join(days_in_week, by = "weekday") %>%
  arrange(weekday_id, daytime) %>%
  ggplot(aes(daytime, weekday_id, fill = mean_weekday_daytime, group = name)) +
  geom_tile(alpha = .5) +
  theme(legend.position = "none") +
  scale_fill_gradient(low = "darkgreen", high = "red") +
  scale_y_reverse(
    breaks = 1:7,
    labels = pull(days_in_week, weekday)
  ) +
  facet_wrap(~name, ncol = 3, scales = "free_y")
```


```{r}
data_background %>%
  select(-stickstoffdioxid, -stickstoffmonoxid) %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name, year, week) %>%
  summarise(mean_year_day = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(week, year, group = name, fill = mean_year_day)) +
  geom_tile(alpha = .7) +
  # scale_x_datetime(date_breaks = "month") +
  # scale_fill_gradient(low = 'darkgreen', high = 'red') +
  scale_fill_viridis_c() +
  scale_x_continuous(breaks = seq(0, 52, 5)) +
  scale_y_reverse(breaks = unique(pull(data_background, year)))
```


```{r}
data_background %>%
  select(-stickstoffdioxid, -stickstoffmonoxid) %>%
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name, year, yearday) %>%
  summarise(mean_year_yearday = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(yearday, year, group = name, fill = mean_year_yearday)) +
  geom_tile(alpha = .7) +
  # scale_x_datetime(date_breaks = "month") +
  # scale_fill_gradient(low = 'darkgreen', high = 'red') +
  scale_fill_viridis_c() +
  scale_x_continuous(
    breaks = seq(0, 365, 10),
    guide = guide_axis(n.dodge = 2)
  ) +
  scale_y_reverse(breaks = unique(pull(data_background, year))) +
  labs(fill = "mean")
```

Compare this years week 12 and 13 to those from the past
```{r}
data_background %>% 
  select(-stickoxide) %>% 
  filter(week %in% c(12, 13)) %>% 
  pivot_longer(cols = one_of(variables)) %>%
  group_by(name, year) %>%
  summarise(mean_year_kw1213 = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(year, mean_year_kw1213)) +
  geom_col(aes(fill = mean_year_kw1213)) +
  scale_x_continuous(labels = as.integer) +
  scale_fill_gradient(low = "darkgreen", high = "red") +
  facet_wrap(~name, ncol = 1)
```

